### 2012-10-29 paper_sean/hyperneat.tex
lots o editing

\subsubsection{HyperNEAT.}

\editbox{JBY: this whole section could be shortened if needed. At the very least, it could be rewritten to seem less like it was ripped out of another paper. Work on this last.}

HyperNEAT is an algorithm for evolving artificial neural networks (ANNs)~\cite{stanley2009hypercube}. It has been repeatedly described in detail~\cite{stanley2009hypercube,gauci2007generating,clune2011performance},  so here we provide only a summary. Instead of directly encoding the weights of the ANN individually on the genome, in HyperNEAT the genome is a compositional pattern producing network (CPPN)~\cite{stanley2007CPPN}. The CPPN genome specifies the weights in a similar way to how natural organisms develop. In nature, phenotypic attributes are specified as a function of their geometric location, and such positional information is conveyed through chemical morphogen gradients~\cite{carroll2005endless}. For example, the concentration of one chemical could indicate the position along the anterior-posterior (head-to-tail) axis, and another chemical in bands could indicate if a cell is in an odd-numbered or even-numbered segment. Based on the relative concentrations of these chemicals, a cell can know where it is geometrically and, thus, what type of cell to become~\cite{carroll2005endless}.

With CPPNs this process is abstracted as a network of mathematical functions that operate in a Cartesian geometric space. The coordinates of phenotypic elements are provided as inputs to the CPPN and the output specifies a phenotypic trait. For example, when CPPNs encode 2D pictures, the coordinates of each pixel are iteratively input into the genome and the output is the grayscale value at that coordinate~\cite{secretan2011picbreeder}. Because a CPPN network is composed of math functions, those functions can create geometric regularities in the phenotype. For example, a Gaussian function of the x-axis can provide left-right symmetry, and a repeating function (e.g. sine) of the y-axis could provide a related motif such as segmentation~(\figref{hyperneat_bug_example}). Both 2D pictures and 3D objects evolved with CPPNs look like natural and engineered objects, and contain complex, regularities like symmetries and repeated motifs, with and without variation~\cite{secretan2011picbreeder,clune2011objects}.

@@

%%% FIGURES %%%

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=hyperneat.eps, width=8cm}
% \caption [ ]{Producing ANNs from CPPNs. In HyperNEAT, weights are specified as a function of the Cartesian coordinates for source and input nodes and a constant bias for each connection between the source and input. All pairwise combinations of source and target node coordinates are iteratively passed into CPPNs to determine the weight of each ANN link. Figure taken from Clune et al.~\cite{clune2}}.
% \end{center}
% \end{figure}
\figp{hyperneatExplanation}{1}{A CPPN specifying a neural network. In HyperNEAT, weights are a function of the Cartesian coordinates of the source and target node for each connection. All pairwise combinations of source and target node coordinates are iteratively passed into a CPPN to determine the weight of each ANN link. Figure from Clune et al. 2011~\cite{clune2011performance}.}

>>>>

\subsubsection{HyperNEAT:}

HyperNEAT is an algorithm for evolving artificial neural networks (ANNs)~\cite{stanley2009hypercube}. It has been repeatedly described in detail~\cite{stanley2009hypercube,gauci2007generating,clune2011performance},  so here we provide only a summary. Instead of directly encoding the weights of the ANN individually on the genome, in HyperNEAT the genome is a compositional pattern producing network (CPPN)~\cite{stanley2007CPPN}. The CPPN genome specifies the weights in a similar way to how natural organisms develop. In nature, phenotypic attributes are specified as a function of their geometric location, and such positional information is conveyed through chemical morphogen gradients~\cite{carroll2005endless}. For example, the concentration of one chemical could indicate the position along the anterior-posterior (head-to-tail) axis, and another chemical in bands could indicate if a cell is in an odd- or even-numbered segment. Based on the relative concentrations of these chemicals, a cell can know where it is geometrically and, thus, what type of cell to become~\cite{carroll2005endless}.

With CPPNs this process is abstracted as a network of math functions that operate in a Cartesian geometric space. The coordinates of phenotypic elements are provided as inputs to the CPPN and the output specifies a phenotypic trait. For example, when CPPNs encode 2D pictures, the coordinates of each pixel are iteratively input into the genome and the output is the grayscale value at that coordinate~\cite{secretan2011picbreeder}. Because a CPPN network is composed of math functions, those functions can create geometric regularities in the phenotype. For example, a Gaussian function of the x-axis can provide left-right symmetry, and a repeating function (e.g. sine) of the y-axis could provide a related motif such as segmentation~(\figref{hyperneat_bug_example}). Both 2D pictures and 3D objects evolved with CPPNs look like natural and engineered objects, and contain complex, regularities such as symmetries and repeated motifs, with and without variation~\cite{secretan2011picbreeder,clune2011objects}.

@@

\figp{hyperneatExplanation}{1}{A CPPN specifying a neural network. In HyperNEAT, weights are a function of the Cartesian coordinates of the source and target node for each connection. All pairwise combinations of source and target node coordinates are iteratively passed into a CPPN to determine the weight of each ANN link. Figure from Clune et al. 2011~\cite{clune2011performance}.}

### 2012-10-29 paper_sean/results.tex
lots o editing

All learning in this study took place in simulation, but the ultimate
goal of our evolved gaits was to perform well in reality. Thus, we
compare our learning performance both in simulation and in hardware
(reality) to previous studies where data is available.

\subsubsection{Simulation}

\edit{Punt for now, update this once we have results from Kyrre} 
The gaits evolved with HyperNEAT in simulation
significantly outperformed the gaits evolved with a GA by Glette et
al. 2012~(Table~\ref{results}). We compared Kyrre et al 2012's
best simulated run to the average of our 20 runs and found that our
gait is, in average, greater in fitness by 7.67 cm/s, with 2.0 cm/s
standard deviation, for every generation of evolution during the
runs. Using the same data, we recorded a fitness of 21.0 cm/s with
4.29 cm/s standard deviation as the mean fitness throughout the
evolutionary process while Kyrre et al 2012's data recorded a fitness
of 13.35 cm/s with 4.06 cm/s standard deviation. These results show
that in simulation, HyperNEAT's gaits evolve faster and have higher
fitnesses than those of GA's.

Also add: This result
is even more significant when one considers that the gait produced by
HyperNEAT with a simulator evolved with 20000 (33\%) fewer evaluations
per run than Glette et al. 2012, a decision mandated by the available
computational resources.

\subsubsection{Hardware}

Gaits evolved in simulation in this study and then transferred to a
physical robot outperformed gaits from all previous studies that were
evolved using only hardware evaluations~\cite{yos:clune,haocheng}
(\tabref{results}). The first important result is the confirmation of
the hypothesis proposed earlier: that using a simulator, although
imperfect, can benefit the evolution of gaits for the real world.

The most likely explanation for this benefit is the large difference
in the population size afforded by a simulator (200 for simulation
vs. 9 in hardware) and the number of generations (200 vs. 20), leading
to a total difference in the number of evaluations between the two
studies of 40000 vs 180 per evolutionary run~\cite{yos:clune}. Another
potential cause of improved performance is that with a simulator
HyperNEAT had much less noise in the evaluation process during early
generations: it could thus find coordinated, regular, gaits, which are
both higher-performing and more likely to transfer \edit{Jeff, what do you think?}. On the physical
robot, the noise in the evaluation was substantial, preventing
effective learning and the discovery of effective
gaits~\cite{yos:clune}. Moreover, we cannot rule out that the
addition in this study of a punishment for high-frequency gaits aided
performance, especially since in simulation the gaits were
high-frequency in a few early generations before rapidly settling to a
lower frequency due to the penalty~(\figref{avg_fit_vs_gen},
right). It is possible that quickly leaving the high-frequency area of
the search space caused a significant performance boost.

In addition to showing the benefits of using a simulator, the second
important result is that HyperNEAT is able to find faster real world gaits
than a simpler encoding~\cite{glette} when both are evolved in simulation.

However, comparing performance in hardware between studies performed in
different laboratories is difficult, not only because reality is
inherently noisy, but because even two copies of the same robot are
not identical and may produce different gait speeds for the same input
gait. In particular, one copy of the QuadraTot robot ---
CCML\footnote{Cornell Creative Machines Lab} --- was used to evaluate
gaits for this study and previous studies~\cite{yos:clune,haocheng},
but a different copy of the same robot --- ROBIN\footnote{Robotics and
Intelligent Systems, U. Oslo} --- was used for another
study~\cite{glette}.

The previous fastest gait on any copy of a QuadraTot was from Glette
et al. 2012~\cite{glette} at 17.8 cm/s, as measured on ROBIN. The
fastest gait measured in our study on our robot, CCML, was 14.5
cm/s. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from \cite{glette} 10 times on CCML and measured an average speed of
only 12.95 cm/s (vs. 17.8 cm/s as measured on ROBIN) and a maximum of
13.76 cm/s (\tabref{results}). Thus, it appears that our robot,
CCML, is slightly slower than ROBIN.

Further, we note that all learning for both this paper
and \cite{glette} was done in simulation, without any influence from
either hardware copy of the robot. Because gaits produced in this
study performed better than all previously recorded gaits on CCML (the only
platform on which they have been compared), we tentatively
conclude that this higher performance would generalize to new copies
of the QuadraTot platform.

\subsubsection{Discussion}

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & 16.4       &   13.0   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\end{center}
\tablabel{results}
\caption{Gait speeds in simulation and on each of two QuadraTot hardware platforms for this paper and three previous studies. We report the number of evaluations per run, simulated (average) and real trial (max) velocities in cm/s.}  %\tablabel{results}
\end{table}

Qualitatively, the HyperNEAT gaits are regular and coordinated, as has
been found in previous simulation
studies~\cite{clune2009evolving,clune2011performance}. This is evident
in the servo position plots where the gaits produced in this study are
smooth and are similar between the servos~(see \figref{servo_plot_111}
for a representative example). This result is important because it
confirms that HyperNEAT can produce the important property of
regularity in simulation, which was not previously observed when
evolving directly on the hardware~\cite{yos:clune}.

While this study was able to produce the fastest QuadraTot gait to
date, most of the gaits in simulation did not transfer well to
reality. Many gaits that performed very well in simulation performed
poorly on the real robot, largely due to servos that were too weak and
timed out, or because of frictional differences between simulation and
reality. Repeated attempts to minimize these problems were
unsuccessful.  A major, unavoidable cause of the servo shutdowns was
the design of the robot: the force required to lift the robot in many
poses was too much for the servos. Future studies may use a
robot that has more mechanical advantage and requires less torque from
each servo, such as the Aracna platform~\cite{lohmann2012aracna},
which was designed in response to the challenges experienced with the
QuadraTot platform.

\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}
{Performance plots. \textbf{Left: }Fitnesses of the champion gaits
averaged over 20 runs, measured in distance travelled over 12 seconds
in cm/s. These gaits outperformed gaits evolved in simulation using a
genetic algorithm~\cite{glette} by 54.3\% in simulation and 5.4\% in
reality, and they outperformed gaits evolved using HyperNEAT in
hardware by 49.5\% (\tabref{results}). \edit{Add Kyrre's runs (+ Haocheng?), switch
from STD to SE if necessary.}\textbf{Right: }Average frequencies of
the gaits averaged over 20 runs. After implementing a policy which
punished gaits with high frequencies to reduce servo shutdowns,
HyperNEAT was able to find high-performing, low frequency
gaits. \edit{add line at 1.67 Hz}}

\figp{servo_plot_111}{0.95}
{Servo positions over time (for nine servos) for a gait produced in
this study. The smoothness, symmetry, and regularities of the gait
show that by using a simulator, HyperNEAT was able to fully exploit
the geometry of the problem.}

>>>>

%We
%compare performance in simulation and hardware
%to available data from previous studies.
%\subsubsection{Simulation:}

The simulator enabled HyperNEAT to evolve
fast gaits that look like those of natural animals. In simulation, HyperNEAT gaits were significantly faster than those from Glette et al. (\figref{FINAL_kyrre_sean_fitness}, $\emph{p} < 6.8\times10^{-8}$, comparing the best gaits in the final generation of each run via MATLAB's Wilcoxon rank sum test).  
Specifically, HyperNEAT gaits were 55\% faster on average in simulation (25.4
cm/s $\pm$ 3.4 SD versus 16.7 cm/s $\pm$ 1.9 SD, \tabref{results}).\edit{Table 1, not 3} 
Plots of servo positions over time reveal that the evolved HyperNEAT gaits are regular, smooth and coordinated~(\figref{servo_plot_111}), confirming previous results with HyperNEAT in simulation~\cite{clune2009evolving,clune2011performance}.

\figp{FINAL_kyrre_sean_fitness}{0.62}
{HyperNEAT outperforms a genetic algorithm with a simpler encoding, when both algorithms are combined with a simulator. Plotted are means over 20 runs in simulation (solid lines) $\pm$ STD (dashed lines). HyperNEAT gaits are 54.3\% faster in
simulation and 5.4\% faster in reality~(\tabref{results}).}

\figgp{servo_plot_111}{.48}{frequencyPlotThreshFinal}{.48}
{\textbf{Left: }
Servo positions over time (for nine servos) for a simulated HyperNEAT gait. HyperNEAT produces gaits that are smooth, symmetrical, and contain other regularities. 
\textbf{Right: }Mean gait frequency averaged over 20 runs. 
Gaits with frequencies above a threshold (horizontal line) receive a fitness penalty. HyperNEAT quickly learns to produce gaits with frequencies low enough to avoid this penalty. }

On the physical robot, HyperNEAT gaits from this study outperformed gaits from all previous Quadratot studies~(\tabref{results})~\cite{yos:clune,haocheng,glette}, including those of Glette et al. 
 However, comparing performance in hardware between studies performed in
different laboratories is difficult, not only because reality is
inherently noisy, but because even two copies of the same robot are
not identical and may produce different gait speeds for the same input
gait. Gaits for this study and two previous studies~\cite{yos:clune,haocheng} were evaluated on a copy of the QuadraTot robot in the Cornell Creative Machines Lab (CCML), while the gaits in Glette et al. were evaluated on a different copy of the same robot in the Robotics and
Intelligent Systems (ROBIN) lab at the University of Oslo.

The previous fastest gait on any copy of a QuadraTot was from Glette
et al., and traveled 17.8 cm/s on the ROBIN QuadraTot. The
fastest gait produced by HyperNEAT with a simulator in the experiments for this paper traveled 14.5
cm/s on the CCML QuadraTot. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from Glette et al. 10 times on CCML and measured an average speed of
only 12.95 cm/s (vs. 17.8 cm/s measured on ROBIN) and a maximum of
13.76 cm/s (\tabref{results}). It thus appears that the CCML version of the robot is slower. Moreover, the HyperNEAT gait (14.5 cm/s) is faster than the best gait of Glette et al. on the CCML QuadraTot. Unfortunately, it was not possible to test the best HyperNEAT gait on the ROBIN QuadraTot. Because HyperNEAT outperformed the simple encoding from Glette et al. on the same robotic hardware, we tentatively
conclude that HyperNEAT produces faster gaits for the QuadraTot robot. This conclusion is supported by the fact that HyperNEAT outperformed the encoding from Glette et al. in simulation.

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.0   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\end{center}
\tablabel{results}
\caption{Velocities of learned gaits in simulation and on two different copies of the QuadraTot robot. Subject to availability, data are reported from the experiments for this paper and three previous studies. Reported are the total number of evaluations per run, the average of the fastest gaits produced in each run in simulation, and the fastest gait produced on a physical robot, either the QuadraTot in the Cornell Creative Machines Lab (CCML) or the QuadraTot in the Robotics and Intelligent Systems (ROBIN) lab at the University of Oslo. *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those 20 runs. Velocities are in cm/s, and bold indicates the best performance.}  
\end{table}

We now discuss the differences between the gaits evolved by HyperNEAT directly on the physical robot, and those first evolved in a simulator and then transferred to the robot. 
On the physical robot, the gaits produced by HyperNEAT with a simulator were faster, more natural, and more repeatable than those evolved directly on the Quadratot robot~\cite{yos:clune}. The gaits were also more regular, as they were in simulation~(\figref{servo_plot_111}). This result is important because it confirms that HyperNEAT can produce the important property of
regularity in a challenging, real-world domain, which was not previously observed when
evolving directly on the hardware~\cite{yos:clune}. Producing regular solutions is a key to exploiting regularity in difficult engineering problems~\cite{clune2011performance}.

The simulator likely improved performance because of the number of evaluations it enabled, both in terms of the population (200 vs. 9) and the number of generations (200 vs. 20), leading
to a total difference of 40000 vs 180 per evolutionary run~\cite{yos:clune}. Another
potential cause of improved performance is that with a simulator
HyperNEAT had much less noise in the evaluation process during early
generations and could thus find coordinated, regular, gaits, which are
both higher-performing and more likely to transfer. On the physical robot, the noise in the evaluation was
substantial, preventing effective learning and the discovery of
effective gaits~\cite{yos:clune}. It could also be the case that 
the addition in this study of a punishment for high-frequency gaits
aided performance, especially since in simulation the gaits were
high-frequency in a few early generations before rapidly settling to a
frequency range below the penalized threshold~(\figref{servo_plot_111},
right).

While this study was able to produce the fastest QuadraTot gait to
date, most of the gaits in simulation did not transfer well to
reality. Many gaits that were fast in simulation performed
poorly on the real robot, largely due to servos that were too weak and
shut down, or because of differences between simulation and
reality. Repeated attempts to minimize these problems were
unsuccessful.  A major, unavoidable cause of the servo shutdowns was
the design of the robot: the force required to lift the robot in many
poses was too much for the servos. Future studies will use a
robot that has more mechanical advantage and requires less torque from
each servo, such as the Aracna platform~\cite{lohmann2012aracna},
which was designed in response to the challenges experienced with the
QuadraTot platform.

### 2012-10-29 paper_sean/robot.tex
lots o editing

% !TEX root =main.tex
\subsubsection{Robot Hardware.}
%\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{captioncaption}

We performed the experiments on the QuadraTot quadrupedal robot platform~(\figref{quadratotWhiteBooties}).
It has 9 degrees-of-freedom (DOF): two joints per leg and one joint that allowed rotation along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
The QuadraTot hardware designs and software for this project are open source\footnote{A parts list, hardware CAD files, software (including the simulator), and gait videos are available at http://creativemachines.cornell.edu/evolved-quadruped-gaits}, and all hardware components are either off-the-shelf or 3D printed parts. %These traits enable other labs to perform research using the same platform, which eases comparison of different gait-learning algorithms. 
To date there are results on the platform from nine different learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}.

Both in this work and that of Glette et al. 2012~\cite{glette}, the joints are powered by Robotis Dynamixel servos; five AX-18A servos for the inner joint of each leg and the single midline joint, and four AX-12A servos for the the outer joint of each leg, which requires less power and can thus have less expensive motors. The servos were sent new positions at 40Hz via the Pydynamixel library. Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque was too high~\cite{robotis}. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in a previous study~\cite{yos:clune}, gaits generated during evolution are highly variable and generate many shutdowns because they forced the servos to exert too much torque or command positions that are out of range. %JMC: Does requesting a value out of range really cause the servo to shut down? That seems like something we could easily fix (by never allowing an out-of-range value). Now that we do not allow values out of range, we still see lots of shutdowns...making me think that the problem was really torque. What do you think?
To prevent collisions between different pieces of the robot's body, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. 2012~\cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al. 2011~\cite{yos:clune} had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable.

We tracked the robot's position using an infrared LED on the robot and tracked it with a Wiimote.

>>>>

% !TEX root =main.tex
\subsubsection{Robot Hardware:}
%\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{captioncaption}

We performed the experiments on the QuadraTot quadrupedal robot platform~(\figref{quadratotWhiteBooties})~\cite{yos:clune}.
It has 9 degrees of freedom: two joints per leg and one joint that allows rotation along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
The QuadraTot hardware designs and the software for this project are open source\footnote{A parts list, hardware CAD files, software (including the simulator), and gait videos are available at http://creativemachines.cornell.edu/evolved-quadruped-gaits}, and all hardware components are either off-the-shelf or 3D printed parts. %These traits enable other labs to perform research using the same platform, which eases comparison of different gait-learning algorithms. 
To date there are results on the platform for nine different learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}.

The joints are powered by Robotis Dynamixel servos; five AX-18A servos for the inner joint of each leg and the single midline joint, and four AX-12A servos for the the outer joint of each leg, which requires less power and can thus have less expensive motors. The servos were sent new positions at 40Hz via the Pydynamixel library. Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque is too high. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in a previous study~\cite{yos:clune}, gaits generated during evolution on QuadraTot are highly variable and produce many shutdowns because they force the servos to exert too much torque.
To prevent collisions between different pieces of the robot's body, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. 2012~\cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al. 2011 had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable. We tracked the robot's position using an infrared LED observed by a Wiimote.
